---
title: "Model USA Biogeoclimatic Units using Machine Learning"
author: "William H MacKenzie and Kiri Daust"
date: "17/10/2019"
output: html_document
---
# Script to create RF model from USA training points and predict + map US zones/ subzones (predicted within each zone)
Import training data
Reduce variable list with caret or PCA all variable
Build machine learning model
Test for missing variable space or map error for required additional training points
Loop to build models by zone and then by subzone within zone to predict map
Translates predictions to a grid point map
Add predictions to hex polygon layer.
Overlay additional plots over map to assign to USA_BGC units


# STEP 1: Prepare dataset for analysis----clear workspace
```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(rattle)
require(rpart)
require(rpart.plot)
require(plyr)
require(reshape)
require(reshape2)
#require(VSURF)
#require(reports)
#require(rpart.utils)
#require(rfUtilities)
require(parallel)
require(foreach)
require(doParallel)
require(ggplot2)
#require(functional)
require(plot3D)
require(dplyr)
require(tcltk)
require(caret)
require(randomForest)
require(ranger)
#require (OpenMP)
#require (randomForestSRC)
require (tools)
require(data.table)
require(spatstat)
require(spatialEco)
require(survey)
require(scales)
require(UBL)
#require (mcl3)
require(tidyr)
require (tidyverse)
require(rlang)
require (Rcpp)
require(sf)
require (mapview)
require(forcats)
require(Boruta)
require(purrrlyr)
require(skimr)
require(gbm)
require(vcd)
require(alluvial)
#install.packages ("alluvial")
```

# Import USA training point data 
CSV file with all variable climate data generated by ClimateNA for the 1960-91 Normal Period.
Several additional climate variables are calculated
```{r input data, echo=FALSE}

X1 <- fread("./inputs/training_pts/US_TrainingPointsAll_31Mar2020_Normal_1961_1990MSY.csv",  stringsAsFactors = FALSE,data.table = FALSE)
X1 <- X1 %>% dplyr::select(-BGC)
## update BGC from new table with updated BGC
X1_new <- fread("./inputs/training_pts/US_TrainingPointsAll_15May2020_dm.csv",  stringsAsFactors = FALSE,data.table = FALSE)
X1_new <- X1_new %>% dplyr::select(ID1, BGC)
##temporary updates
# X1_new2 <- fread("./inputs/training_pts/CWHvm_to_CWHms_2.csv",  stringsAsFactors = FALSE,data.table = FALSE)
# X1_new2$BGC <- recode(X1_new2$BGC, "CWHvm_WA" = "CWHms_WA") 
# X1_new2 <- X1_new2 %>% dplyr::select(ID1, BGC) 
# X1_new <- left_join(X1_new, X1_new2, by = "ID1")
# X1_new$BGC.y <- ifelse(is.na(X1_new$BGC.y), X1_new$BGC.x, X1_new$BGC.y)
# X1_new <- X1_new %>% select(ID1, BGC.y) %>% rename(BGC = BGC.y)
## update BGC from new table with updated BGC

X1 <- left_join(X1, X1_new, by = "ID1")

#X1 <- fread(file.choose(),  stringsAsFactors = FALSE,data.table = FALSE)

X1_loc <- X1 %>% dplyr::select(ID1, Latitude, Longitude, Elevation)
X1 <- X1 %>% dplyr::select (ID1, BGC, everything()) %>% dplyr::select(-Latitude, -Longitude, -Elevation)

# X2 <- fread("./cLHS/WNA_Training_22_VARS.csv", stringsAsFactors = FALSE, data.table = FALSE)
# X3 <- left_join(X2, X1) %>% dplyr::select(ID1, BGC, Latitude, Longitude, Elevation)
# write.csv(X3, "LHS_Training_Locations.csv")

addVars <- function(dat){
  dat$PPT_MJ <- dat$PPT05 + dat$PPT06  # MaY/June precip
  dat$PPT_JAS <- dat$PPT07 + dat$PPT08 + dat$PPT09  # July/Aug/Sept precip
  dat$PPT.dormant <- dat$PPT_at + dat$PPT_wt  # for calculating spring deficit
  dat$CMD.def <- 500 - (dat$PPT.dormant)  # start of growing season deficit original value was 400 but 500 seems better
  dat$CMD.def[dat$CMD.def < 0] <- 0  #negative values set to zero = no deficit
  dat$CMDMax <- dat$CMD07
  dat$CMD.total <- dat$CMD.def + dat$CMD
  dat$CMD.grow <- dat$CMD05 + dat$CMD06 +dat$CMD07 +dat$CMD08 +dat$CMD09
  dat$DD5.grow <- dat$DD5_05 + dat$DD5_06 + dat$DD5_07 + dat$DD5_08 + dat$DD5_09
  dat$CMDMax <- dat$CMD07 # add in so not removed below
  dat$DDgood <- dat$DD5 - dat$DD18
  dat$DDnew <- (dat$DD5_05 + dat$DD5_06 +dat$DD5_07  + dat$DD5_08)  - (dat$DD18_05 + dat$DD18_06 +dat$DD18_07 +dat$DD18_08)
  dat$TmaxJuly <- dat$Tmax07
  dat$tmaxSum20 <-  dat %>%
      select( starts_with("Tmax")) %>% -20 %>%
              mutate_all(., list(~ifelse(. < 0, 0 , .))) %>%
                    rowSums(na.rm = TRUE)

dat$tmaxSum25 <-  dat %>%
      select( starts_with("Tmax")) %>% -25 %>%
              mutate_all(., list(~ifelse(. < 0, 0 , .))) %>%
                    rowSums(na.rm = TRUE)

dat$tmaxSum30 <-  dat %>%
      select( starts_with("Tmax")) %>% -30 %>%
              mutate_all(., list(~ifelse(. < 0, 0 , .))) %>%
                    rowSums(na.rm = TRUE)

dat$tmaxSum35 <-  dat %>%
      select( starts_with("Tmax")) %>% -35 %>%
              mutate_all(., list(~ifelse(. < 0, 0 , .))) %>%
                    rowSums(na.rm = TRUE)

# remove some redundant variables considered undesireable
month <- c("01", "02", "03", "04", "05", "06","07", "08", "09", "10", "11", "12")
dat <- dat  %>% dplyr::select(-ends_with(month)) %>% #removes all monthly variables
  dplyr::select(-starts_with("Rad")) %>% ##remove other non-biological variables
  dplyr::select(-starts_with("RH")) %>%
  dplyr::select (-contains("DD_0")) %>%
   dplyr::select  (-contains("DD18")) %>%
   dplyr::select  (-contains("DD_18"))  %>%
  dplyr::select( -PPT_sp, -FFP, -PAS_sm, -PPT_at, -PAS_at, - MAP, -TD, -MAT)
  return(dat)
}


X1 <- addVars(X1)
#$BGC <- as.factor(X1$BGC)
```
The preprocessing function from the caret package was used to identify variables with near-zero variance or correlation >0.90 in the combined data set. These variables were removed leaving a final variable set of 20 variables.
```{r reduce variables, warning=FALSE} 
########Remove near zero variance and highly correlated variables
X1_no_nzv_pca <- preProcess(X1[,-c(1:2)], method = c( "nzv")) # DROP variables with near zero variance
X1_no_nzv_pca
X1 <- dplyr::select(X1, -c(X1_no_nzv_pca$method$remove))
X1 <- X1 %>%  na_if(-9999.0) %>% drop_na()

# calculate correlation matrix of variables
correlationMatrix <- cor(X1[,-c(1:2)])
# summarize the correlation matrix
#print(correlationMatrix)
# find attributes that are highly corrected (ideally >0.75)
highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.95, verbose = TRUE, names = TRUE) ## review for removal of highly correlated vars
X1 <- X1 %>% dplyr::select(-c(highlyCorrelated))%>% drop_na(BGC)


modelvars <- colnames(X1 %>% dplyr::select(-ID1,  -BGC, everything()))
X1 <- X1 %>% dplyr::select(ID1, BGC, all_of(modelvars)) 

# modelvars <- read.csv("./inputs/Final14Var_WNABGCv11.csv")##use model variables from buildrf work
# X1 <- X1 %>% dplyr::select(ID1, BGC, modelvars$x) %>% drop_na(BGC)
X1$BGC <- as.factor(X1$BGC)
X1 <- droplevels(X1)

```

# remove undersampled BGCs
```{r remove undersampled}
#removeBGCs <- c("ICHdm", "IDFxx1")
#X2 <- X1 %>% filter(!(BGC %in% removeBGCs)) %>% droplevels()
# merge parkland into adjacent alpine unit
X1$BGC <- X1$BGC %>% recode( MHdsp_OR = "CMAun_OR", MHRFdmp_OR = "CMAun_OR", MHRFmmp_OR = "CMAun_OR", MHmsp_WA = "CMAun_WA",
                             "MHRFdsp_CA" = "CMAun_CA", "MHdsp_OR" = "CMAun_OR" , "MHRFdmp_OR"= "CMAun_OR" , "MHRFmmp_OR"= "CMAun_OR" , 
                            "MHmsp_WA" ="CMAun_WA" , "MHmmp_WA"= "CMAun_WA" ,
                             "ESSFxcp_CO"=  "IMAun_CO" , "ESSFwmp_MT"="IMAun_MT" , "ESSFxkp_UT"= "IMAun_UT" , "ESSFxkp_MT"= "IMAun_MT" , 
                             "ESSFdmp_ID"="IMAun_ID" ,"ESSFdkp_MT"=  "IMAun_MT" ,"ESSFxcp_WA" = "IMAun_WA" ,"ESSFmwp_WA" = "IMAun_WA" ,
                           "ESSFxwp_OR" = "IMAun_OR" )
## merge of related units with high confusion in first run
#X1$BGC <- X1$BGC %>% recode( CWHvm_OR = "CWHvm_WA", MHRFdmp_OR = "CMAun_OR")
#                             , MHRFmmp_OR = "CMAun_OR", MHmsp_WA = "CMAun_WA",
#                              "MHRFdsp_CA" = "CMAun_CA", "MHdsp_OR" = "CMAun_OR" , "MHRFdmp_OR"= "CMAun_OR" , "MHRFmmp_OR"= "CMAun_OR" , 
#                             "MHmsp_WA" ="CMAun_WA" , "MHmmp_WA"= "CMAun_WA" ,
#                              "ESSFxcp_CO"=  "IMAun_CO" , "ESSFwmp_MT"="IMAun_MT" , "ESSFxkp_UT"= "IMAun_UT" , "ESSFxkp_MT"= "IMAun_MT" , 
#                              "ESSFdmp_ID"="IMAun_ID" ,"ESSFdkp_MT"=  "IMAun_MT" ,"ESSFxcp_WA" = "IMAun_WA" ,"ESSFmwp_WA" = "IMAun_WA" ,
#                            "ESSFxwp_OR" = "IMAun_OR" )
 BGCs <- unique(X1$BGC)
BGCs
#X1 <- X2
```

##Remove outliers points
Looks for training points which fall climatically well outside of the normal range by BGC.
Mostly aimed at catching the 10% of FIA plots which have been given an intentionally large offset georeferenced location
May also indicate overly broad BGCs (eg. IMAus)

```{r remove outlier points}

# X1_Bio1 <- X1_Bio1  %>% select( -BGC) %>% select (Zone, everything())
# # require(psych)
# # require(outliers)
# # X1_test <- X1_Bio1[X1_Bio1$Zone == "CMA",]
# # X1_test <- X1_test[,-c(4:17)]
#  outlier <- psych::outlier(X1_Bio1[-1])
#  X1_Bio1 [,-c(1)] <- dmap (X1_Bio1 [,-c(1)], as.numeric)

 

removeOutlier <- function(dat, alpha,numIDvars){
  out <- foreach(curr = unique(as.character(dat$BGC)), .combine = rbind) %do% {
    temp <- dat[dat$BGC == curr,]
    md <- tryCatch(mahalanobis(temp[,-c(1:numIDvars)],
                               center = colMeans(temp[,-c(1:numIDvars)]), 
                               cov = cov(temp[,-c(1:numIDvars)])), error = function(e) e)
    if(!inherits(md,"error")){
      ctf <- qchisq(1-alpha, df = ncol(temp)-1)
      outl <- which(md > ctf)
      cat("Removing", length(outl), "outliers from",curr, "; ")
      if(length(outl) > 0){
        temp <- temp[-outl,]
      }
    }
    temp
    
  }
  return(out)
}

 X2 <- removeOutlier(X1, alpha = .001,numIDvars = 2)

 ####show histogram of variables
 skimmed <- skim(X2)
skimmed



featurePlot(x =  X1[, -c(1:2)], 
            y =  X2$Zone, 
            plot = "box",
            strip=strip.custom(par.strip.text=list(cex=.7)),
            scales = list(x = list(relation="free"), 
                          y = list(relation="free")))

```

# Build model of BGCs
The model reduced via RFE to 14 variables is approximately 84.3% accurate and 83% Gini. When all variables are used (44) the accuracy and Gini increase by approximately 1%.
Stay with limited set

```{r build ranger model training point size, cache = TRUE, echo=FALSE}

# set.seed(123321)
# coreNo <- makePSOCKcluster(detectCores() - 1)
# registerDoParallel(coreNo, cores = detectCores() - 1)
# Cores <- as.numeric(detectCores()-1)
# ctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 5,
#                      classProbs = FALSE, verboseIter = T,
#                      savePredictions = "final", 
#                      allowParallel = F)
# 
# 
# #BGCmodel_rang <- train(Zone  ~ ., data = X1_Bio1[-2], #for zone
# BGCmodel_subz <- train(BGC  ~ ., data = X1[,-c(1)], # for subzone
#                      method = "ranger")#,
#                      #preProcess = c("center", "scale", "YeoJohnson"), #,tuneGrid = tgrid,
#                       trControl = ctrl,
#                       importance = "impurity")#, splitrule = "extratrees")
# 
#  stopCluster(coreNo)
#  gc()
# BGCmodel_subz
# # predict the outcome on a test set
# BGC_pred <- predict(BGCmodel_subz , X1[,-c(1:2)])
# # compare predicted outcome and true outcome
# con <- confusionMatrix(BGC_pred, as.factor(X1$BGC)) 
# BGCmodel <- BGCmodel_subz
# file=paste("./outputs/USA_SubZone_RFModel_Mar11",".Rdata",sep="")
# save(BGCmodel,file=file)


#############simple ranger model
BGCmodel <- ranger(BGC ~ ., data = X2[,-c(1)], #do.trace = 10,
                           num.trees = 501,   importance = "impurity", splitrule = "extratrees", seed = 12345, write.forest = TRUE, classification = TRUE)#strata=BGC, sampsize= c(500),, probability = TRUE

BGCmodel
#save(BGCmodel,file= "./outputs/USAv11_26_VAR_SubZone_ranger.Rdata")
v <-as.data.frame(BGCmodel$variable.importance)
DF <- v %>% tibble::rownames_to_column() %>% rename(w = rowname, v = 'BGCmodel$variable.importance')
 
 ggplot(DF, aes(x=reorder(w,v), y=v,fill=v))+ 
  geom_bar(stat="identity", position="dodge")+ coord_flip() +
   ylab("Variable Importance")+
   xlab("")+
   ggtitle("Information Value Summary")+
   guides(fill=F)+
   scale_fill_gradient(low="red", high="blue")

 fname <- "USA_28var"
model = "_ranger_10May2020"
write.csv (BGCmodel$variable.importance, file= paste("./outputs/",fname,"_Importance",model,".csv",sep=""))
write.csv (BGCmodel$prediction.error, file= paste("./outputs/",fname,"_Error",model,".csv",sep=""))
write.csv (BGCmodel$confusion.matrix, file= paste("./outputs/",fname,"_ConfusionMatrix",model,".csv",sep=""))

####alternate model build
 fname <- "USA_28var"
model = "_rf_13May2020_2"
BGCmodel_rf <- randomForest(BGC ~ ., data = X2[,-c(1)], do.trace = 10,
                          ntree = 101,   importance = TRUE)#strata=BGC, sampsize= c(500),, probability = TRUE

write.csv (BGCmodel_rf$confusion, file= paste("./outputs/",fname,"_ConfusionMatrix",model,".csv",sep=""))
write.csv (BGCmodel_rf$err.rate, file= paste("./outputs/",fname,"_error",model,".csv",sep=""))
```

```{r build points file of misclassified point}
###Predict
X2$BGC.pred <- predict(BGCmodel_rf, data = X2[,-c(1:2)])

X2.mis <- X2 %>% select(ID1, BGC, BGC.pred) %>% filter(BGC != BGC.pred) %>% left_join(X1_loc)

write.csv(X2.mis, "./outputs/Confused_Points09_May2020.csv", row.names = FALSE)

```


```{r graphical confusion for ranger}
# xtab <- BGCmodel_subz$pred
# cm <- confusionMatrix(data = xtab$pred, reference = xtab$obs)
# cm.tab <- as.matrix(cm)
# cm.ovr <- as.matrix(cm,what="overall")
# cm.cls <- as.matrix(cm,what="classes")
# write.csv(cm.cls, "./outputs/SubzoneConfusionStatistics.csv")
# mosaic(cm$table)
# plotCM <- function(cm){
#   cmdf <- as.data.frame(cm[["table"]])
#   cmdf[["color"]] <- ifelse(cmdf[[1]] == cmdf[[2]], "green", "red")
#   
#   alluvial::alluvial(cmdf[,1:2]
#                      , freq = cmdf$Freq
#                      , col = cmdf[["color"]]
#                      , alpha = 0.5
#                      , hide  = cmdf$Freq == 0
#                      )
# }
# plotCM(cm)

```

####Predict BGC membership of hex grid
### Use the best model from those above (Biological, nzv and corr removed, Scaled, Outliers Removed)
``` {r Build }
#load("./outputs/USA_Zone_RFModel_Dec4.Rdata")
# bring in grid data with climate attributes
Z1 <- fread("./inputs/grids/USA_800m_HexPts_Normal_1961_1990MSY.csv", stringsAsFactors = FALSE,  data.table = FALSE)
Z1_loc <-  Z1 %>% dplyr::select(ID1, Latitude, Longitude, Elevation)
#grid.dat <- fread("./inputs/grids/USA_8000m_Grid_Climate_new_Normal_1961_1990MSY.csv", stringsAsFactors = FALSE,  data.table = FALSE)### add in the extra climate variables
Z1$PPT_MJ <- Z1$PPT05 + Z1$PPT06 # MaY/June precip
Z1$PPT_JAS <- Z1$PPT07 + Z1$PPT08 + Z1$PPT09 # July/Aug/Sept precip
Z1$PPT.dormant <- Z1$PPT_at + Z1$PPT_wt # for calculating spring deficit
Z1$CMD.def <- 500 - (Z1$PPT.dormant)# start of growing season deficit original value was 400 but 500 seems better
Z1$CMD.def [Z1$CMD.def < 0] <- 0 #negative values set to zero = no deficit
Z1$CMD.grow <- Z1$CMD05 + Z1$CMD06 +Z1$CMD07 +Z1$CMD08 +Z1$CMD09
Z1$DD5.grow <- Z1$DD5_05 + Z1$DD5_06 + Z1$DD5_07 + Z1$DD5_08 + Z1$DD5_09
Z1$CMD.total <- Z1$CMD.def + Z1$CMD
Z1$CMDMax <- Z1$CMD07 # add in so not removed below
Z1$DDgood <- Z1$DD5 - Z1$DD18
Z1$DDnew <- (Z1$DD5_05 + Z1$DD5_06 +Z1$DD5_07  + Z1$DD5_08)  - (Z1$DD18_05 + Z1$DD18_06 +Z1$DD18_07 +Z1$DD18_08)
Z1$TmaxJuly <- Z1$Tmax07
modelvars <- read.csv("./inputs/Final22Var_WNABGCv11.csv")##use model variables from buildrf work
modelvars <- modelvars %>% filter(x !="BGC")
grid.dat <- Z1 %>% dplyr::select(ID1,modelvars$x) %>%  filter_all(all_vars(. > -9999))

###Predict
grid.pred <- predict(BGCmodel, data = grid.dat[,-c(1)])

Z1_loc <- Z1_loc[Z1_loc$ID1 %in% grid.dat$ID1,]
Z1.pred <- cbind(Z1_loc, grid.pred$predictions)
#Z1.
#grid.dat$Zone <- as.factor(grid.dat$Zone)
Z1.pred$BGC <-  fct_explicit_na(Z1.pred$`grid.pred$predictions` , na_level = "(None)")
#write.csv(grid.dat, "./outputs/Hexgrid800mPredictedtoSubZone_30vars.csv", row.names = FALSE)
Z1.pred <- dplyr::select(Z1.pred, ID1, BGC, Latitude, Longitude,Elevation )#"Longitude","Latitude", "Elevation", "Zone")]
write.csv(Z1.pred, "./outputs/Hexgrid800mPredictedtoSubZone_May5.csv", row.names = FALSE)
```
# Attribute hex grid with Zone call
```{r link to hex polygon layer}
##############link predicted Zones to Polygons and write shape file
hexpoly <- st_read(dsn = "./inputs/hex_shapes", layer = "USA_bgc_hex_800m")
hexZone <- left_join(hexpoly, Z1.pred, by = c("FID" = "ID1"))
temp <- select(hexZone, BGC, geometry)
temp2 <- st_zm(temp, drop=T, what='ZM')
st_write(temp2, dsn = "outputs", layer = "USA_SubZoneMap_800m_31Mar2020_ungrouped", driver = "ESRI Shapefile", update = TRUE)

# #####joined polygons have bad topology. INSTEAD DISSOLVE in QGIS.
# temp2 <- temp
# st_precision(temp2) <- 0.5 ###union polygons within zone
# temp2$BGC <-  as.factor(temp2$BGC)
# temp2$BGC <- forcats::fct_explicit_na(temp2$BGC,na_level = "(None)")
# droplevels(temp2$BGC)
# t2 <- temp2 %>%
#   group_by(BGC) %>%
#   summarise(geometry = sf::st_union(geometry)) %>%
#   ungroup()
# #mapView(t2)
# t2 <- st_zm(t2, drop=T, what='ZM')
# st_write(t2, dsn = "outputs", layer = "USA_rawSubZoneMap_800m_ranger_31Mar2020", driver = "ESRI Shapefile", update = TRUE)

```
## clean up isolated individual pixels
This process is slow. IS probably better done in QGIS.
In QGIS
Take the multipolygon, separate into polygons (v fast), multipart to single parts, clip to state boundary, dissolve with v.dissolve, add geometric attributes to (v fast), select polygons <2 000 000 = 3 hexes in size (v fast), eliminate selected polygons  by longest neighbour.

``` {r cleanup up crumbs}

###now cleanup and remove crumbs
# t2 <- st_cast(t2, "MULTIPOLYGON") %>% st_cast("POLYGON")
# t2 <- t2 %>%
#   mutate(Area = st_area(.)) %>%
#   mutate(ID = seq_along(Zone))
# unique(t2$Area)
# 
# 
# library(units)
# size <- 2000000
# size <- set_units(size, "m^2")
# tSmall <- t2[t2$Area <= size,]
# t2$Zone <- as.character(t2$Zone)
# 
# require(doParallel)
# coreNum <- as.numeric(detectCores()-1)
# coreNo <- makeCluster(coreNum)
# registerDoParallel(coreNo, cores = coreNum)
# 
# ###loop through each polygon < size, determine intersects, and assign to zone with most edge touching
# ###all the built in functions I found only dealt with holes in the middle of polygons
# new <- foreach(i = 1:length(tSmall$ID), .combine = rbind, .packages = c("foreach","sf")) %dopar% {
#   ID <- tSmall$ID[i]
#   nbrs <- st_intersects(tSmall[i,],t2)[[1]]
#   nbrs <- nbrs[!nbrs %in% ID]
#   if(length(nbrs) == 0){return(NULL)}
#   lines <- st_intersection(t2[ID,],t2[nbrs,])
#   lines <- st_cast(lines)
#   l.len <- st_length(lines)
#   names(l.len) <- lines$Zone.1
#   zn <- names(l.len)[l.len == max(l.len)][1]
#   newDat <- t2[ID,]
#   newDat$Zone <- zn
#   newDat
# }
# 
# stopCluster(coreNo)
# gc()
# temp <- t2[!t2$ID %in% new$ID,]
# t2 <- rbind(temp, new) %>%
#   mutate(Zone = as.factor(Zone))
# 
# ###now have to combine crumbs with existing large polygons
# temp2 <- t2
# st_precision(temp2) <- 0.5
# t2 <- temp2 %>%
#   group_by(Zone) %>%
#   summarise(geometry = sf::st_union(geometry)) %>%
#   ungroup()
# 
# mapview(t2, zcol = "Zone")
# t2 <- st_zm(t2, drop=T, what='ZM')
# st_write(t2, dsn = "outputs", layer = "DeCrumbed_SubZoneMap_800m_ranger30", driver = "ESRI Shapefile", update = TRUE)

```


# Build Subzone models and predict within each Zone
####now build models for each subzone and then predict gridpoints by subzone model
Advice of Tom Hengel is that single model is more mathematically justified

``` {r Build Subzone model and predict}

# # Grid data from previous chunk
# pointDat <- grid.dat
# pointDat <-  pointDat %>% select(Zone, everything()) %>% arrange(Zone)
# 
# ######################################################
# #Training point data from above
# X2 <-  X1.info %>% select(Zone, everything()) %>% arrange(Zone)
# X2$BGC <-  as.factor(X2$BGC)
# Zones <- as.character(unique(X2$Zone))
# Zones <- sort (Zones)
# X2 <- droplevels(X2)
# # X2 <- X2[X2$Zone == "CMA",]
# # Zones <- as.character(unique(X2$Zone))
# # Zones <- sort (Zones)
# SZPred <- foreach(Z = Zones, .combine = rbind) %do% {
#   trainSub <- X2[X2$Zone == Z,]###subset training points to only include selected zone
#    if(length(unique(trainSub$BGC)) >= 2){ ###somtimes there aren't any subzones skip to else
#     trainSub$BGC <- as.factor(trainSub$BGC)
#     trainSub$BGC <- droplevels(trainSub$BGC)
#     trainSub <- removeOutlier(trainSub, alpha = 0.001)
# ###build model for each subzone individually
#     set.seed(123321)
# coreNo <- makePSOCKcluster(detectCores() - 1)
# registerDoParallel(coreNo, cores = detectCores() - 1)
# Cores <- as.numeric(detectCores()-1)
# 
# SZmodel <- train(BGC  ~ ., data = trainSub [-1],
#                      method = "ranger",
#                      trControl = trainControl(method="cv", number = 10, verboseIter = T, classProbs = T, savePredictions = "final"),
#                      #num.trees = 11,
#                      preProcess = c("center", "scale", "YeoJohnson"),#,tuneGrid = tgrid,
#                      importance = "impurity")
# 
#  stopCluster(coreNo)
# gc()
#       # SZmodel <- randomForest(BGC ~ ., data=trainSub [-1], nodesize = 5, do.trace = 10,
#       #                       ntree=101, na.action=na.fail, importance=TRUE, proximity=FALSE)
# 
#    
#     pointSub <- pointDat[pointDat$Zone == Z,] ###subset grid based on previously predicted zones
#     pointSub$BGC<- predict(SZmodel, newdata = pointSub[,-c(1:4),]) ### predict subzones
#     out <- pointSub[,c("ID1", "Latitude","Longitude","Zone", "BGC")]
#     out
#   }else{ ##if only one subzone, plot
#     pointSub <- pointDat[pointDat$Zone == Z,]
#     #pointSub$Zone <- droplevels(pointSub$Zone)
#     pointSub$BGC <- pointSub$Zone
#     out <- pointSub[,c("ID1", "Latitude","Longitude","Zone", "BGC")]
#     out
#   }
#   
# }
# 
# grid.sbz <- dplyr::select(SZPred, ID1, BGC, Zone)#"Longitude","Latitude", "Elevation", "Zone")]
# grid.sbz <- droplevels(grid.sbz)
# table(grid.sbz$BGC)
# grid.dat.pred <- left_join(grid.dat, grid.sbz, by = c("ID1", "Zone"))
# grid.dat.pred <- grid.dat.pred %>% select(ID1, Zone, BGC, everything())
# write.csv(grid.dat, "./outputs/Hexgrid800mPredictedtoSubZone_ranger.csv", row.names = FALSE)
# ```
# # Attribute hexpolygons with subzone predictions 
# ```{r link subzone predictions to hex polygons}
# ###link to hex polygons
# hexZone2 <- left_join(hexZone, grid.sbz, by = "ID1")
# #temp <- select(hexZone, Zone, geometry)
# temp <- hexZone2
# temp$BGC <-  fct_explicit_na(temp$BGC , na_level = "(None)")
# table(temp$BGC)
# 
# temp2 <- temp
# st_precision(temp2) <- 0.5###union polygons within zsubone
# t3 <- temp2 %>%
#   group_by(BGC) %>%
#   summarise(geometry = sf::st_union(geometry)) %>%
#   ungroup()
# 
# mapview(t3)
# t3 <- st_zm(t3, drop=T, what='ZM')
# st_write(t3, dsn = "outputs", layer = "SubzoneRaw800m_ranger_30", driver = "ESRI Shapefile", update = TRUE)


```

## This should probably be done from the spatial file since it will have be simplified with decrumbing

```{r cLHS of trainingpoints for WNA}
#select a subset of training points by BGC from the final surface for full model build
# library(clhs)
# #BGC = "ICHxwz"
# #countZone <- points %>% count(Zone)
# countSubzone <- grid.dat.pred %>% count(BGC)
# 
# rownames(grid.dat.pred) <- grid.dat.pred[,1]
# countSubzone$logn <- log(countSubzone$n, 10)
# countSubzone$rs <- as.integer (rescale(countSubzone$logn, to = c(500, 1200), from = range(countSubzone$logn, na.rm = TRUE, finite = TRUE)))
# countSubzone$sample <- ifelse(countSubzone$rs > countSubzone$n, countSubzone$n, countSubzone$rs )
# write.csv (countSubzone, "./outputs/USA_pts_per_BECLHC2.csv")
# 
# allUnits <- unique(grid.dat.pred$BGC)
# #grid.dat.pred$ID1 <- row.names(grid.dat.pred)
# set.seed(123321)
# coreNo <- makePSOCKcluster(detectCores() - 1)
# registerDoParallel(coreNo, cores = detectCores() - 1)
# Cores <- as.numeric(detectCores()-1)
# 
# BGC = "BGdh_OR"
# LHCtraining <- foreach(BGC = allUnits, .combine = rbind, .packages = c("clhs")) %dopar% {
# temp <- grid.dat.pred[(grid.dat.pred$BGC %in% BGC),]
# temp_names <- temp[1:6]
# Num <- countSubzone$sample[(countSubzone$BGC %in% BGC)]
#   samples <- clhs(temp[,-c(1:6)], 
#                 size = Num,           # Test a range of sample sizes
#                 iter = 10000,        # Arbitrarily large number of iterations of optimization procedure. Default=10,000 but a larger number may be used
#                 progress = TRUE, 
#                 simple = FALSE)
# 
# cLHS <- samples$sampled_data
# cLHS$ID1 <- row.names (cLHS)
# cLHS_Points <- merge (temp[,c(1:5)], cLHS, by = "ID1")
# cLHS_Points
# }
#  stopCluster(coreNo)
#  gc()
# 
# write.csv(LHCtraining, "./outputs/USA_Training_LHC_w_data.csv", row.names = FALSE)
# LHCtraining2 <- LHCtraining [1:3] 
# rownames(grid.dat.raw) <- grid.dat.raw[,1]
# LHCtraining2 <- left_join(LHCtraining2, grid.dat.raw, by = "ID1")
# write.csv(LHCtraining2, "./outputs/USA_LHS_all_dat.csv", row.names = FALSE)
# USA_LHS <- dplyr::select(LHCtraining2, ID1, BGC, Longitude, Latitude, Elevation)
# write.csv(USA_LHS, "./outputs/USA_LHS_for_ClimateNA.csv", row.names = FALSE)
#X1 <- LHCtraining #feed these back into building a new rF model above
```
