---
title: "Model USA Biogeoclimatic Units using Machine Learning"
author: "William H MacKenzie and Kiri Daust"
date: "17/10/2019"
output: html_document
---
# Script to create RF model from USA training points and predict + map US zones/ subzones (predicted within each zone)
Import training data
Reduce variable list with caret or PCA all variable
Build machine learning model
Test for missing variable space or map error for required additional training points
Loop to build models by zone and then by subzone within zone to predict map
Translates predictions to a grid point map
Add predictions to hex polygon layer.
Overlay additional plots over map to assign to USA_BGC units


# STEP 1: Prepare dataset for analysis----clear workspace
```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(rattle)
require(rpart)
require(rpart.plot)
require(plyr)
require(reshape)
require(reshape2)
#require(VSURF)
#require(reports)
#require(rpart.utils)
#require(rfUtilities)
require(parallel)
require(foreach)
require(doParallel)
require(ggplot2)
#require(functional)
require(plot3D)
require(dplyr)
require(tcltk)
require(caret)
require(randomForest)
require(ranger)
#require (OpenMP)
#require (randomForestSRC)
require (tools)
require(data.table)
require(spatstat)
require(spatialEco)
require(survey)
require(scales)
require(UBL)
#require (mcl3)
require(tidyr)
require (tidyverse)
require(rlang)
require (Rcpp)
require(sf)
require (mapview)

#install.packages ("haven")
```

# Import training point data 
CSV file with all variable climate data generated by ClimateNA for the 1960-91 Normal Period
```{r input data, echo=FALSE}

X1 <- fread("inputs/training_pts/US_TrainingPointsAll_21Nov2019_renamed_Normal_1961_1990MSY.csv",  stringsAsFactors = FALSE,data.table = FALSE)
## add some additional calculated variables
X1$PPT_MJ <- X1$PPT05 + X1$PPT06 # MaY/June precip
X1$PPT_JAS <- X1$PPT07 + X1$PPT08 + X1$PPT09 # July/Aug/Sept precip
X1$PPT.dormant <- X1$PPT_at + X1$PPT_wt # for calculating spring deficit
X1$CMD.def <- 500 - (X1$PPT.dormant)# start of growing season deficit original value was 400 but 500 seems better
X1$CMD.def [X1$CMD.def < 0] <- 0 #negative values set to zero = no deficit
X1$CMDMax <- X1$CMD07
X1$CMD.grow <- X1$CMD05 + X1$CMD06 +X1$CMD07 +X1$CMD08 +X1$CMD09
X1$CMD.total <- X1$CMD.def + X1$CMD

X1$tmaxSum20 <-  X1 %>%
      select( starts_with("Tmax")) %>% -20 %>%
              mutate_all(., list(~ifelse(. < 0, 0 , .))) %>%
                    rowSums(na.rm = TRUE)

X1$tmaxSum25 <-  X1 %>%
      select( starts_with("Tmax")) %>% -25 %>%
              mutate_all(., list(~ifelse(. < 0, 0 , .))) %>%
                    rowSums(na.rm = TRUE)

X1$tmaxSum30 <-  X1 %>%
      select( starts_with("Tmax")) %>% -30 %>%
              mutate_all(., list(~ifelse(. < 0, 0 , .))) %>%
                    rowSums(na.rm = TRUE)

X1$tmaxSum35 <-  X1 %>% 
      select( starts_with("Tmax")) %>% -35 %>%
              mutate_all(., list(~ifelse(. < 0, 0 , .))) %>% 
                    rowSums(na.rm = TRUE)

X1save = X1
X1 = X1save

```


#Reduce the variable space

```{r reduce variable set, echo=FALSE}
#### BioAll model - All Biologically Interpretable Variables Only:
model = "BioAll"

TEMP.list=c("Tmax_sp","Tmax_sm","Tmin_wt","Tmin_sp","Tmin_sm","Tmin_at",
            "DD5_sp","DD5_sm","DD5_at","DD5", "Tmax07","Tmax08",
            "MAT","MWMT","MCMT", "EMT","EXT", "DD5_06", "DD5_05", "TD","tmaxSum30")
#            "tmaxSum20", "tmaxSum25", ,"tmaxSum35" )
PPT.list=c("PPT06", "PPT_sp", "PPT_sm","MSP", "MAP","PAS", "PPT_MJ", "PPT_JAS")
OTHER.list=c("CMD_sp","CMD_sm","AHM","SHM","NFFD","bFFP","FFP","NFFD_sp", "CMD", "CMD.grow", "CMD07", "CMD.def", "CMD.total", "Eref")
ClimateVar=c(TEMP.list,PPT.list,OTHER.list)
List=c("BGC")
X1save = X1
X1$BGC  <- as.factor(X1$BGC)
X1=X1[,names(X1) %in% c(List,ClimateVar)]
X1 <-  drop_na (X1)
#X1 <- X1[! X1$BGC == NA,]
X1$BGC <- as.factor(X1$BGC)
X1 <- droplevels(X1)
X1.sub <- X1
#X1.sub$BGC <- as.character(X1.sub$BGC)


#X1.sub$BGC <- gsub("_","",X1.sub$BGC) ## for subzone
X1.sub$BGC <- as.factor(X1.sub$BGC)
X1.sub <- droplevels(X1.sub)
X1.sub$BGC <- factor(as.character(X1.sub$BGC))

##identify near zero variance variables
nzv <- nearZeroVar(X1.sub, saveMetrics= TRUE)
nzv[nzv$nzv,][1:10,]
##remove these variables
  nzv <- nearZeroVar(X1.sub)
  X1.sub <- X1.sub[, -nzv]

# calculate correlation matrix of variables
correlationMatrix <- cor(X1.sub[,-c(1)])
# summarize the correlation matrix
print(correlationMatrix)
# find attributes that are highly corrected (ideally >0.75)
highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.95, verbose = TRUE, names = TRUE) ## review for removal of highly correlated vars

X1 <- X1.sub[, !(colnames(X1.sub) %in% highlyCorrelated), drop = FALSE]
X1.var <- colnames(X1)

```
# Build model of Zones
``` {r build model of Zones}
####for equal number of training points
left = function(text, num_char) {
  substr(text, 1, num_char)
}
X1$Zone <- left(X1$BGC, 4)
X1$Zone <- gsub("[[:lower:]]|[[:space:]]|[[:digit:]]", "", X1$Zone) ###for zone model
X1 <- X1[X1$Zone != "",]
X1$Zone <- as.factor(X1$Zone)

tpNum <- 750 ##number of desired training points per unit
mList <- list()
for(Zone in unique(X1$Zone)){##Calculate multiplier
  num <- length(X1$Zone[X1$Zone == Zone])
  mNum <- tpNum/num
  mList[[Zone]] <- mNum
}

###scaled by log10
mList <- list()
for(BGC in unique(X1$Zone)){##Calculate multiplier
  num <- length(X1$Zone[X1$Zone == Zone])
  tpNum <- log10(num)*300
  mNum <- tpNum/num
  mList[[Zone]] <- mNum
}

###SMOTE!
X1.smote <- SmoteClassif(Zone ~ ., X1[-1], C.perc = mList, k= 5 , repl = TRUE, dist = "Euclidean")
X1.smote <- X1[-1]
table(X1.smote$BGC)

BGCmodel <- randomForest(Zone ~ ., data=X1.smote, nodesize = 5, do.trace = 10,
                         ntree=101, na.action=na.fail, importance=TRUE, proximity=FALSE)

```

####Predict zone membership of hex grid
``` {r Build }


grid.dat <- fread("./inputs/grids/USA_8000m_Grid_Climate_new_Normal_1961_1990MSY.csv", stringsAsFactors = FALSE,  data.table = FALSE)
Z1 <- grid.dat
## add some additional calculated variables
Z1$PPT_MJ <- Z1$PPT05 + Z1$PPT06 # MaY/June precip
Z1$PPT_JAS <- Z1$PPT07 + Z1$PPT08 + Z1$PPT09 # July/Aug/Sept precip
Z1$PPT.dormant <- Z1$PPT_at + Z1$PPT_wt # for calculating spring deficit
Z1$CMD.def <- 500 - (Z1$PPT.dormant)# start of growing season deficit original value was 400 but 500 seems better
Z1$CMD.def [Z1$CMD.def < 0] <- 0 #negative values set to zero = no deficit
Z1$CMDMax <- Z1$CMD07
Z1$CMD.grow <- Z1$CMD05 + Z1$CMD06 +Z1$CMD07 +Z1$CMD08 +Z1$CMD09
Z1$CMD.total <- Z1$CMD.def + Z1$CMD

Z1$tmaxSum20 <-  Z1 %>%
      select( starts_with("Tmax")) %>% -20 %>%
              mutate_all(., list(~ifelse(. < 0, 0 , .))) %>%
                    rowSums(na.rm = TRUE)

Z1$tmaxSum25 <-  Z1 %>%
      select( starts_with("Tmax")) %>% -25 %>%
              mutate_all(., list(~ifelse(. < 0, 0 , .))) %>%
                    rowSums(na.rm = TRUE)

Z1$tmaxSum30 <-  Z1 %>%
      select( starts_with("Tmax")) %>% -30 %>%
              mutate_all(., list(~ifelse(. < 0, 0 , .))) %>%
                    rowSums(na.rm = TRUE)

Z1$tmaxSum35 <-  Z1 %>% 
      select( starts_with("Tmax")) %>% -35 %>%
              mutate_all(., list(~ifelse(. < 0, 0 , .))) %>% 
                    rowSums(na.rm = TRUE)

colnames(Z1)[1:2] <- c("ID1","ID2")
# ####New 16 Variable Set
# VarList = c("TD",	"Eref",	"AHM",	"Tmax07",	"CMD.total",	"Tmax_sp","Tmin_sm","CMD_sp","PAS",	
#             "PPT_sp","PPT06","CMD.def","DD5_sp","Tmin_at","bFFP","MCMT")
List = c("ID1",  "Latitude", "Longitude", "Elevation")
grid.dat = Z1[,names(Z1) %in% c(List, X1.var)]
grid.dat2 <- grid.dat
###Predict
grid.dat$Zone <- predict(BGCmodel, newdata = grid.dat[,-c(1:4)])
grid.zone <- dplyr::select(grid.dat, ID1, Zone, Latitude, Longitude,Elevation )#"Longitude","Latitude", "Elevation", "Zone")]

hexpoly <- st_read(dsn = "./inputs/hex_shapes", layer = "USA_HexPoly_8000m_new")
##testpnts <- pointDat[pointDat$Latitude > 43.407 & pointDat$Latitude < 46.5 & pointDat$Longitude < -111.79 & pointDat$Longitude > -115.412,]##subset to small area for example
hexZone <- left_join(hexpoly, grid.zone, by = "ID1")
temp <- select(hexZone, Zone, geometry)
# testpnts <- grid.zone
# testpnts <- testpnts[,c("Longitude","Latitude","Zone")]
# testpnts <- st_as_sf(testpnts, coords = c("Longitude","Latitude"), crs = 4326, agr = "constant")##convert to spatial object
# testpnts <- st_transform(testpnts, crs = st_crs(hexpoly))
# temp <- st_join(hexpoly,testpnts)
# temp <- temp[!is.na(temp$Zone),]
#temp <- temp[,-1]

temp2 <- temp
st_precision(temp2) <- 0.5###union polygons within zone
t1 <- temp2 %>%
  group_by(Zone) %>%
  summarise(geometry = sf::st_union(geometry)) %>%
  ungroup()
mapView(t1)

st_write(t1, dsn = "outputs", layer = "rawZoneMap", driver = "ESRI Shapefile", update = TRUE)

```



# Build Subzone models and predict within each Zone

``` {r Build Subzone model and predict}
####now subzones
#t1 <- st_read(dsn = "outputs", layer = "rawZoneMap")
pointDat <- grid.dat
######################################################

X2 <-  X1
X2 <- X1 %>%  filter(Zone == "ICH" | Zone == "BG")
X2 <- X1 %>%  filter(Zone == "BG")
Zones <- as.character(unique(X2$Zone))
#Zones <- as.character(unique(pointDat$Zone))
#Zones <- sort (Zones)

#X2 <- droplevels(X2)

SZPred <- foreach(Z = Zones, .combine = rbind) %do% {
  trainSub <- X2[X2$Zone == Z,]###subset training points to only include selected zone
   #trainSub$BGC <- droplevels(trainSub$Zone)
   if(length(unique(trainSub$BGC)) >= 2){ ###somtimes there aren't any subzones
    trainSub$BGC <- as.factor(trainSub$BGC)
    trainSub$BGC <- droplevels(trainSub$BGC)
    trainSub <- removeOutlier(trainSub, alpha = 0.001)
    mList <- list()
    # for(BGC in unique(trainSub$BGC)){##Calculate multiplier
    #   num <- length(trainSub$BGC[trainSub$BGC == BGC])
    #   tpNum <- log10(num)*150
    #   mNum <- tpNum/num
    #   mList[[BGC]] <- mNum
    # }
    trainSub <- SmoteClassif(BGC ~ ., trainSub [-19], C.perc = mList, k= 5 , repl = TRUE, dist = "Euclidean")
    
 # drop = "Zone"
      SZmodel <- randomForest(BGC ~ ., data=trainSub [-19], nodesize = 5, do.trace = 10, 
                            ntree=101, na.action=na.fail, importance=TRUE, proximity=FALSE)
    
    pointSub <- pointDat[pointDat$Zone == Z,] ###subset grid based on previously predicted zones
    pointSub$BGC<- predict(SZmodel, newdata = pointSub[,-c(1:4,22),]) ### predict subzones
    
    out <- pointSub[,c("ID1", "Latitude","Longitude","Zone", "BGC")]
    out
  }else{ ##if only one subzone, plot 
    pointSub <- pointDat[pointDat$Zone == Z,]
    #pointSub$Zone <- droplevels(pointSub$Zone)
    pointSub$BGC <- pointSub$Zone
    out <- pointSub[,c("ID1", "Latitude","Longitude","Zone", "BGC")]
    out
  }
  
}

grid.sbz <- dplyr::select(SZPred, ID1, BGC)#"Longitude","Latitude", "Elevation", "Zone")]

hexZone <- left_join(hexZone, grid.sbz, by = "ID1")
#temp <- select(hexZone, Zone, geometry)
temp <- hexZone
temp$BGC <-  fct_explicit_na(temp$BGC , na_level = "(None)")
table(temp$BGC)

# grid.sbz <- SZPred[,c("Longitude","Latitude","BGC")]
# grid.sbz  <- st_as_sf(testpnts, coords = c("Longitude","Latitude"), crs = 4326, agr = "constant")##convert to spatial object
# grid.sbz  <- st_transform(grid.sbz , crs = st_crs(hxPnts))
# temp <- st_join(hxPnts,grid.sbz )
# temp <- temp[!is.na(temp$BGC),]
# temp <- temp[,-1]

temp2 <- temp
st_precision(temp2) <- 0.5###union polygons within zone
t3 <- temp2 %>%
  group_by(BGC) %>%
  summarise(geometry = sf::st_union(geometry)) %>%
  ungroup()
st_write(t3, dsn = "outputs", layer = "SubzoneRaw", driver = "ESRI Shapefile", update = TRUE)
mapview(t3)

# hexZone <- left_join(hexpoly, grid.sbz , by = "ID1")
# temp <- select(hexZone, BGC, geometry)
# # testpnts <- grid.zone
# # testpnts <- testpnts[,c("Longitude","Latitude","Zone")]
# # testpnts <- st_as_sf(testpnts, coords = c("Longitude","Latitude"), crs = 4326, agr = "constant")##convert to spatial object
# # testpnts <- st_transform(testpnts, crs = st_crs(hexpoly))
# # temp <- st_join(hexpoly,testpnts)
# # temp <- temp[!is.na(temp$Zone),]
# #temp <- temp[,-1]
# 
# temp2 <- temp
# st_precision(temp2) <- 0.5###union polygons within zone
# t3 <- temp2 %>%
#   group_by(Zone) %>%
#   summarise(geometry = sf::st_union(geometry)) %>%
#   ungroup()
# mapView(t3)
# 
# st_write(t3, dsn = "outputs", layer = "rawZoneMap", driver = "ESRI Shapefile", update = TRUE)

```

##Build machine learning model

```{r plots of variables}

require(doParallel)
set.seed(123321)
coreNo <- makePSOCKcluster(detectCores() - 1)
registerDoParallel(coreNo, cores = detectCores() - 1)
Cores <- as.numeric(detectCores()-1)
 control <- trainControl(method = 'cv', number = 2,  allowParallel = TRUE)
   BGCmodel <- train(BGC ~ ., data = X1.sub,  method ='rf', trControl = control, ntree = 21)
                     #, do.trace = 10, ntree = 11)
                         # ntree=71, mtry=3, nodesize = 5, metric = metric, na.action=na.omit, importance=FALSE, proximity=FALSE)#strata=BGC, sampsize= c(500),
 stopCluster(coreNo)
 gc()
  print(BGCmodel)
  
  rf.imp <- varImp(BGCmodel, scale = FALSE)
rf.imp
  ##may need to change BGC for Zone where running for entire model of for subzone by zone foreach loop
# ptm <- proc.time()
# BGCmodel <- ranger(BGC ~ ., data=X1, verbose = TRUE, 
#                         num.trees=71, importance="impurity")
# proc.time() - ptm
fit.rf <- randomForest(BGC ~ ., data=X1.sub2,  do.trace = 10, mtry= 3,
                       ntree=71, na.action=na.omit, importance=TRUE, proximity=FALSE)

 varImpPlot(fit.rf, sort=TRUE, n=40)  

    ###caret package version
Seed <- 1234
metric <- "Kappa"
 control <- trainControl(method = 'repeatedcv', number = 2, repeats = 2, allowParallel = TRUE)
  BGCmodel <- train(BGC ~ ., data=X1.sub,  method='rf', trControl = control, do.trace = 10, 
                          ntree=71, mtry=3, nodesize = 5, metric = metric, na.action=na.omit, importance=FALSE, proximity=FALSE)#strata=BGC, sampsize= c(500),

 #metric <- "Accuracy"

# Save Confusion matrix
print(BGCmodel$confusion, digits=2)
#Novel <-outlier(BGCmodel, X1)
#fname <- "BGCv11_AB_USA_500each_27VAR_SubZone"
fname <- "BC_var16"
model = "Oct22"
write.csv(BGCmodel$proximity, file= paste(fname,"_Proximity",model,".csv",sep=""))
write.csv(BGCmodel$importance, file= paste(fname,"_Importance",model,".csv",sep=""))
write.csv(BGCmodel$err.rate, file= paste(fname,"_Error",model,".csv",sep=""))
write.csv(BGCmodel$confusion, file= paste(fname,"_ConfusionMatrix",model,".csv",sep=""))
write.csv(BGCmodel$confusion[, 'class.error'], file= paste(fname,"_Confusion",model,".csv",sep=""))
VIP <- varImpPlot(BGCmodel, sort=TRUE) 
write.csv(VIP, file= paste(fname,"_OBO",model,".csv",sep=""))
dev.copy(pdf,(paste(model,'VarImpPlot.pdf')))
dev.off()


###BGCmodel from ranger implementation of randomForest *****not much faster
ptm <- proc.time()
BGCmodel_ran <- ranger(BGC ~ ., data=X1,  mtry=NULL, write.forest = TRUE, num.trees=11, importance='impurity', classification = TRUE)
proc.time() - ptm

#### Save random forest model
file=paste(fname,"_RFmodelv4",".Rdata",sep="")
save(BGCmodel,file=file)

# Save Confusion matrix
print(BGCmodel$confusion, digits=2)
write.csv(BGCmodel$err.rate, file= paste(fname,"_ErrorBio2v1000_53",".csv",sep=""))
write.csv(BGCmodel$confusion, file= paste(fname,"_ConfusionMatrixBio2v1000_53",".csv",sep=""))
write.csv(BGCmodel$confusion[, 'class.error'], file= paste(fname,"_ConfusionBio2v1000_53",".csv",sep=""))
varImpPlot(BGCmodel, sort=TRUE, n.var=nrow(BGCmodel$importance))

###Runs iterations to determine how many variables are required for accurancy
ptm <- proc.time()
BGCmodelVar <- rfcv(trainx = X1[,-1], trainy = X1[,1], scale = "log", step=0.5)
proc.time() - ptm
    ###graph of decrease in error with number of variables
with(BGCmodelVar, plot(n.var, error.cv,log="x", type="o", lwd=2))
print(BGCmodelVar$predicted)
save(BGCmodelVar$predicted,file=paste(fname,"_VarImp",".JPEG",sep=""))
      
 ##### run the next line of code iteratively to find best mtry number
BGCmodel2 <- tuneRF(X1[,-1], X1[,1], mtryStart=5, ntreeTry=151, stepFactor=1.5, improve=0.05,trace=TRUE, plot=TRUE, doBest=FALSE)



#stopCluster(coreNo)
##############Parallel version
# ptm <- proc.time()
# 
# Zonemodel <- foreach(ntree = rep(15,7), .combine = randomForest::combine, .multicombine = TRUE, .packages = "randomForest") %dopar% {
#   randomForest(X1[-1], X1[1],
#                ntree=ntree, na.action=na.fail, importance=TRUE, proximity=FALSE)
# }
# proc.time() - ptm
   #### Save random forest model
file=paste("USA_Subzones_RFModel_16Var_20Nov2019",".Rdata",sep="")
save(BGCmodel,file=file)
#load("USA_Subzones_RFModel_16Var_20Nov2019.Rdata")
#load("USA_Subzones_RFModel_27Var.Rdata")
#load(file.choose())

```


#####Predict BGC units##########################
```{r predict BGC for hexgrid}
#load("BGCv11_AB_USA_LHC_10VAR_SubZone_RFmodel.Rdata") ##load random forest model
### Read in US Hex Grid file with normal data. SOMETIMES THE FIRST 2 COLUMN NAMES NEED TO BE CHANGED
Columns = c("rn", "Region", "Latitude", "Longitude", "Elevation", "AHM", "bFFP",
            "CMD07","DD5_sp","EMT","Eref_sm","EXT","FFP","MCMT","MSP",
            "PPT07","PPT08", "PPT05","PPT06","PPT09", "SHM","TD","Tmax_sp","Tmin_at",
            "Tmin_sm","Tmin_wt", "PPT_at","PPT_wt", "PAS","eFFP",
            "Eref09","MAT","Tmin_sp","CMD")#"GCM", 

pointDat <- fread(file.choose(), stringsAsFactors = FALSE,  data.table = FALSE)#, select = Columns
#pointDat <- pointDat[pointDat$DD_0_01 != -9999,]
#pointDat <- pointDat[pointDat$Latitude < 49,]
X1 <-pointDat
### add additional variables to match rF model
X1$PPT_MJ <- X1$PPT05 + X1$PPT06 # MaY/June precip
X1$PPT_JAS <- X1$PPT07 + X1$PPT08 + X1$PPT09 # July/Aug/Sept precip
X1$PPT.dormant <- X1$PPT_at + X1$PPT_wt # for calculating spring deficit
X1$CMD.def <- 500 - (X1$PPT.dormant)# start of growing season deficit original value was 400 but 500 seems better
X1$CMD.def [X1$CMD.def < 0] <- 0 #negative values set to zero = no deficit
X1$CMDMax <- X1$CMD07
X1$CMD.total <- X1$CMD.def + X1$CMD
X1save = X1
X1 = X1save
colnames(X1)[1]=c("ID1")
colnames(X1)[2]=c("ID2")
#####New 16 Variable list to Use.
VarList = c("TD",	"Eref",	"AHM",	"Tmax07",	"CMD.total",	"Tmax_sp","Tmin_sm","CMD_sp","PAS",	
            "PPT_sp","PPT06","CMD.def","DD5_sp","Tmin_at","bFFP","MCMT")#16Var
#VarList = c("AHM", "bFFP","CMD.total","DD5_sp","EMT","Eref_sm","EXT","FFP","MCMT","MSP",
#            "PPT_JAS","PPT_MJ","PPT06","SHM","TD","Tmax_sp","Tmin_at","Tmin_sm","Tmin_wt",
#            "PAS","CMD.def","CMDMax","eFFP","Eref09","MAT","PPT07","Tmin_sp")
List = c("ID1","ID2", "Latitude", "Longitude", "Elevation")
X1.sub = X1
X1.sub=X1[,names(X1) %in% c(List,VarList)]
X1.sub <- X1.sub[!X1.sub[,6] == -9999.0,]
X1 <- X1.sub
pointDat <-X1.sub
#pointDat <- Y1
###Predict BGC units
ranger.pred <- predict(BGCmodel, data = pointDat[,-c(1:5)])
pointDat$BGC.pred <- ranger.pred$predictions
 pointDat.saved <- pointDat
 pointDat <-pointDat.saved
 #points <- pointDat[,c("Latitude","Longitude","Zone")]
 pointDat$BGC.pred <- as.character(pointDat$BGC.pred)
 pointDat$Subzone <- gsub("[[:space:]]","",pointDat$BGC.pred)##for subzones
  pointDat$Zone <- word(pointDat$BGC.pred,1,sep = "\\_")
 pointDat$Zone <- gsub("[[:lower:]]|[[:space:]]|[[:digit:]]","",pointDat$Zone)##for just zone
points <- pointDat[,c("ID1", "Zone", "Subzone", "Latitude","Longitude","Elevation")]
write.csv (points, "TestPredictedPoints.csv")
List2 = c("ID1", "Zone", "Subzone", "Latitude","Longitude","Elevation")
#List2 = c("ID1", "Subzone")
pointLHC <- pointDat[,names(pointDat) %in% c(List2,VarList)]###includes all the climate variables for LHC procedure below
colnames (pointLHC) [18] <- "ID2"
col_order <-c("ID1", "Zone", "Subzone", "Latitude","Longitude","Elevation","CMD.total", "bFFP","Eref_sm","MCMT","PPT_JAS",
              "SHM","TD","PAS","DD5_sp","PPT06")
pointLHC <- pointLHC[,col_order]
#colnames(points)[2]<- "ZoneSubzone"
#colnames(points)[1]<- "ID1"
```

###Write to CSV to import into QGIS and join to HexGrid

``` {r write predicted point data to CSV}
write.csv (points, "USA_FAIPlots_BGCPredicted.csv")
write.csv (pointLHC, "USA_1k_HexGridPts_BGC_predicted_16Var_LHC_May3_w_BGCdata.csv")
countZone <- points %>% count(Zone)
countSubzone <- points %>% count(Subzone)
countSubzone$logn <- log(countSubzone$n, 10)
countSubzone$rs <- as.integer (rescale(countSubzone$logn, to = c(300, 1000), from = range(countSubzone$logn, na.rm = TRUE, finite = TRUE)))
countSubzone$sample <- ifelse(countSubzone$rs > countSubzone$n, countSubzone$n, countSubzone$rs )
allUnits <- unique(points$Subzone)
write.csv (countSubzone, "USA_ptPerBECLHC2.csv")

###########From here add CSV to QGIS and Join by ID1 to HexPolygon Shape file
```  

##############select a subset of training points   
``` {r select a subsample of training points from each predicted BGC for use in WNA model}
library(clhs)
BGC = "ICHxwz"
allUnits <- unique(pointLHC$Subzone)
pointLHC$ID1 <- row.names(pointLHC)
LHCtraining <- foreach(BGC = allUnits, .combine = rbind, .packages = c("clhs")) %dopar% {
temp <- pointLHC[(pointLHC$Subzone %in% BGC),]
Num <- countSubzone$sample[(countSubzone$Subzone %in% BGC)]
  samples <- clhs(temp[,-c(1:6)], 
                size = Num,           # Test a range of sample sizes
                iter = 100,        # Arbitrarily large number of iterations of optimization procedure. Default=10,000 but a larger number may be used
                progress = TRUE, 
                simple = FALSE)

cLHS <- samples$sampled_data
cLHS$ID1 <- row.names (cLHS)
cLHS_Points <- merge (pointLHC[,c(1:6)], cLHS, by = "ID1")
cLHS_Points
}

write.csv(LHCtraining, "USA_Training_LHC.csv")
X1 <- LHCtraining #feed these back into building a new rF model above


