---
title: "Model USA Biogeoclimatic Units using Machine Learning"
author: "William H MacKenzie and Kiri Daust"
date: "17/10/2019"
output: html_document
---
# Script to create RF model from USA training points and predict + map US zones/ subzones (predicted within each zone)
Import training data
Reduce variable list with caret or PCA all variable
Build machine learning model
Test for missing variable space or map error for required additional training points
Loop to build models by zone and then by subzone within zone to predict map
Translates predictions to a grid point map
Add predictions to hex polygon layer.
Overlay additional plots over map to assign to USA_BGC units


# STEP 1: Prepare dataset for analysis----clear workspace
```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(rattle)
require(rpart)
require(rpart.plot)
require(plyr)
require(reshape)
require(reshape2)
#require(VSURF)
#require(reports)
#require(rpart.utils)
#require(rfUtilities)
require(parallel)
require(foreach)
require(doParallel)
require(ggplot2)
#require(functional)
require(plot3D)
require(dplyr)
require(tcltk)
require(caret)
require(randomForest)
require(ranger)
#require (OpenMP)
#require (randomForestSRC)
require (tools)
require(data.table)
require(spatstat)
require(spatialEco)
require(survey)
require(scales)
require(UBL)
#require (mcl3)
require(tidyr)
require (tidyverse)
require(rlang)
require (Rcpp)
require(sf)
require (mapview)

#install.packages ("haven")
```

# Import USA training point data 
CSV file with all variable climate data generated by ClimateNA for the 1960-91 Normal Period.
Several additional climate variables are calculated
```{r input data, echo=FALSE}

X1 <- fread("inputs/training_pts/US_TrainingPointsAll_21Nov2019_renamed_Normal_1961_1990MSY.csv",  stringsAsFactors = FALSE,data.table = FALSE)
## add some additional calculated variables
X1$PPT_MJ <- X1$PPT05 + X1$PPT06 # MaY/June precip
X1$PPT_JAS <- X1$PPT07 + X1$PPT08 + X1$PPT09 # July/Aug/Sept precip
X1$PPT.dormant <- X1$PPT_at + X1$PPT_wt # for calculating spring deficit
X1$CMD.def <- 500 - (X1$PPT.dormant)# start of growing season deficit original value was 400 but 500 seems better
X1$CMD.def [X1$CMD.def < 0] <- 0 #negative values set to zero = no deficit
X1$CMDMax <- X1$CMD07
X1$CMD.grow <- X1$CMD05 + X1$CMD06 +X1$CMD07 +X1$CMD08 +X1$CMD09
X1$CMD.total <- X1$CMD.def + X1$CMD
X1save = X1
X1 = X1save


```
# Various tests to reduce the covariate set
``` {r test the variable importance in caret}
library(caret)
left = function(text, num_char) {
  substr(text, 1, num_char)
}
X1$Zone <- left(X1$BGC, 4)
X1$Zone <- gsub("[[:lower:]]|[[:space:]]|[[:digit:]]", "", X1$Zone) ###for zone model
X1 <- X1[X1$Zone != "",]
X1$Zone <- as.factor(X1$Zone)
X1 <- droplevels(X1)
X1$BGC <- gsub("  ","",X1$BGC) ## for subzone
X1$BGC <- gsub(" ","",X1$BGC) ## for subzone
#### BioAll model - All Biologically Interpretable Variables Only:

TEMP.list=c("Tmax_sp","Tmax_sm","Tmin_wt","Tmin_sp","Tmin_sm","Tmin_at",
            "DD5_sp","DD5_sm","DD5_at","DD5", "Tmax07","Tmax08",
            "MAT","MWMT","MCMT", "EMT","EXT", "DD5_06", "DD5_05", "TD")
PPT.list=c("PPT06", "PPT_sp", "PPT_sm","MSP", "MAP","PAS", "PPT_MJ", "PPT_JAS")
OTHER.list=c("CMD_sp","CMD_sm","AHM","SHM","NFFD","bFFP","FFP","NFFD_sp", "CMD", "CMD.grow", "CMD07", "CMD.def", "CMD.total", "Eref")
ClimateVar=c(TEMP.list,PPT.list,OTHER.list)
List=c("Zone", "BGC")
X1_Bio <- X1[,names(X1) %in% c(List,ClimateVar)]
#########Remove near zero variance and highly correlated variables
X1_Bio_no_nzv_pca <- preProcess(select(X1_Bio, - c(Zone, BGC)), 
                        method = c("nzv","corr"),
                        cutoff = .90)
X1_Bio_no_nzv_pca$method$remove
X1_Bio1 <- select(X1_Bio, -c(X1_Bio_no_nzv_pca$method$remove))
X1_Bio1 <-  drop_na (X1_Bio1)
#X1 <- X1[! X1$BGC == NA,]
X1_Bio1$Zone <- as.factor(X1_Bio1$Zone)
X1_Bio1 <- droplevels(X1_Bio1)
###############remove outliers
# Looks for training points which fall climatically well outside of the normal range by BGC.
# Mostly aimed at catching the 10% of FIA plots which have been given an intentionally offset georeferenced location
# May also indicate overly broad BGCs (eg. IMAus)
X1_Bio1 <-X1_Bio1  %>% select(Zone, everything()) 
X1.var <- colnames(X1_Bio1)

removeOutlier <- function(dat, alpha){
  out <- foreach(curr = unique(as.character(dat$Zone)), .combine = rbind) %do% {
    temp <- dat[dat$Zone == curr,]
    md <- tryCatch(mahalanobis(temp[,-1],center = colMeans(temp[,-1]), cov = cov(temp[,-1])),
                   error = function(e) e
    )
    if(!inherits(md,"error")){
      ctf <- qchisq(1-alpha, df = ncol(temp)-1)
      outl <- which(md > ctf)
      cat("Removing", length(outl), "outliers from",curr, "; ")
      if(length(outl) > 0){
        temp <- temp[-outl,]
      }
    }
    temp
    
  }
  return(out)
}

X1_Bio2 <- removeOutlier(X1_Bio1, alpha = .001)

```

# Build model of Zones

```{r reduce variable set, echo=FALSE}

# Test models between all training points and where outliers removed
set.seed(123321)
coreNo <- makePSOCKcluster(detectCores() - 1)
registerDoParallel(coreNo, cores = detectCores() - 1)
Cores <- as.numeric(detectCores()-1)

BGCmodel2 <- train(Zone  ~ ., data = X1_Bio2[-2],
                     method = "ranger",
                     trControl = trainControl(method="cv", number = 10, verboseIter = T, classProbs = T),
                     #num.trees = 11, 
                     preProcess = c("center", "scale", "YeoJohnson"),#,tuneGrid = tgrid, 
                     importance = "impurity")

 stopCluster(coreNo)
 gc()

BGCmodel <- BGCmodel2
file=paste("./outputs/USA_Zone_RFModel_Nov24",".Rdata",sep="")
save(BGCmodel,file=file)

```

####Predict zone membership of hex grid
### Use the best model from those above (Biological, nzv and corr removed, Scaled, Outliers Removed)
``` {r Build }
#load("./outputs/USA_Zone_RFModel_Nov24.Rdata")
# bring in grid data with climate attributes
grid.dat <- fread("./inputs/grids/USA_800m_Grid_Climate_new_Normal_1961_1990MSY_2.csv", stringsAsFactors = FALSE,  data.table = FALSE)
### add in the extra climate variables
Z1 <- grid.dat
Z1$PPT_MJ <- Z1$PPT05 + Z1$PPT06 # MaY/June precip
Z1$PPT_JAS <- Z1$PPT07 + Z1$PPT08 + Z1$PPT09 # July/Aug/Sept precip
Z1$PPT.dormant <- Z1$PPT_at + Z1$PPT_wt # for calculating spring deficit
Z1$CMD.def <- 500 - (Z1$PPT.dormant)# start of growing season deficit original value was 400 but 500 seems better
Z1$CMD.def [Z1$CMD.def < 0] <- 0 #negative values set to zero = no deficit
Z1$CMDMax <- Z1$CMD07
Z1$CMD.grow <- Z1$CMD05 + Z1$CMD06 +Z1$CMD07 +Z1$CMD08 +Z1$CMD09
Z1$CMD.total <- Z1$CMD.def + Z1$CMD
colnames(Z1)[1:2] <- c("ID1","ID2")
List = c("ID1",  "Latitude", "Longitude", "Elevation")
###reduce variables to list from model selected above
grid.dat = Z1[,names(Z1) %in% c(List, X1.var)]
grid.dat <- grid.dat %>%  filter_all(all_vars(. > -9999))
###Predict
grid.dat$Zone <- predict(BGCmodel, newdata = grid.dat[,-c(1:4)])
grid.zone <- dplyr::select(grid.dat, ID1, Zone, Latitude, Longitude,Elevation )#"Longitude","Latitude", "Elevation", "Zone")]
grid.dat$Zone <-  fct_explicit_na(grid.dat$Zone , na_level = "(None)")
write.csv(grid.dat, "HexgridPredictedtoZone.csv", row.names = FALSE)

##############link predicted Zones to Polygons and write shape file
hexpoly <- st_read(dsn = "./inputs/hex_shapes", layer = "USA_HexPoly_800m_new")
hexZone <- left_join(hexpoly, grid.zone, by = "ID1")
temp <- select(hexZone, Zone, geometry)

temp2 <- temp
st_precision(temp2) <- 0.5###union polygons within zone
t1 <- temp2 %>%
  group_by(Zone) %>%
  summarise(geometry = sf::st_union(geometry)) %>%
  ungroup()
mapView(t1)
t1 <- st_zm(t1, drop=T, what='ZM')
st_write(t1, dsn = "outputs", layer = "rawZoneMap_800m", driver = "ESRI Shapefile", update = TRUE)

```
# clean up isolated individual pixels
This process is slow

``` {r cleanup up crumbs}
# 
# ###now cleanup and remove crumbs
# t2 <- st_cast(t1, "MULTIPOLYGON") %>% st_cast("POLYGON")
# t2 <- t2 %>%
#   mutate(Area = st_area(.)) %>%
#   mutate(ID = seq_along(Zone))
# unique(t2$Area)
# 
# 
# library(units)
# size <- 1555256000#560000
# size <- set_units(size, "m^2")
# tSmall <- t2[t2$Area <= size,]
# t2$Zone <- as.character(t2$Zone)
# 
# require(doParallel)
# coreNum <- as.numeric(detectCores()-1)
# coreNo <- makeCluster(coreNum)
# registerDoParallel(coreNo, cores = coreNum)
# 
# ###loop through each polygon < size, determine intersects, and assign to zone with most edge touching
# ###all the built in functions I found only dealt with holes in the middle of polygons
# new <- foreach(i = 1:length(tSmall$ID), .combine = rbind, .packages = c("foreach","sf")) %dopar% {
#   ID <- tSmall$ID[i]
#   nbrs <- st_intersects(tSmall[i,],t2)[[1]]
#   nbrs <- nbrs[!nbrs %in% ID]
#   if(length(nbrs) == 0){return(NULL)}
#   lines <- st_intersection(t2[ID,],t2[nbrs,])
#   lines <- st_cast(lines)
#   l.len <- st_length(lines)
#   names(l.len) <- lines$Zone.1
#   zn <- names(l.len)[l.len == max(l.len)][1]
#   newDat <- t2[ID,]
#   newDat$Zone <- zn
#   newDat
# }
# 
# stopCluster(coreNo)
# gc()
# temp <- t2[!t2$ID %in% new$ID,]
# t2 <- rbind(temp, new) %>%
#   mutate(Zone = as.factor(Zone))
# 
# ###now have to combine crumbs with existing large polygons
# temp2 <- t2
# st_precision(temp2) <- 0.5
# t2 <- temp2 %>%
#   group_by(Zone) %>%
#   summarise(geometry = sf::st_union(geometry)) %>%
#   ungroup()
# 
# mapview(t2, zcol = "Zone")
# t2 <- st_zm(t2, drop=T, what='ZM')
# st_write(t2, dsn = "outputs", layer = "DeCrumbed_ZoneMap_8000m", driver = "ESRI Shapefile", update = TRUE)

```


# Build Subzone models and predict within each Zone
####now build models for each subzone and then predict gridpoints by subzone model

``` {r Build Subzone model and predict}

# Grid data from previous chunk
pointDat <- grid.dat
pointDat <-  pointDat %>% select(Zone, everything()) %>% arrange(Zone)

######################################################
#Training point data from above
X2 <-  X1_Bio1 %>% select(Zone, everything()) %>% arrange(Zone)
X2$BGC <-  as.factor(X2$BGC)
Zones <- as.character(unique(X2$Zone))
Zones <- sort (Zones)
X2 <- droplevels(X2)

SZPred <- foreach(Z = Zones, .combine = rbind) %do% {
  trainSub <- X2[X2$Zone == Z,]###subset training points to only include selected zone
   if(length(unique(trainSub$BGC)) >= 2){ ###somtimes there aren't any subzones skip to else
    trainSub$BGC <- as.factor(trainSub$BGC)
    trainSub$BGC <- droplevels(trainSub$BGC)
    trainSub <- removeOutlier(trainSub, alpha = 0.001)
###build model for each subzone individually
      SZmodel <- randomForest(BGC ~ ., data=trainSub [-1], nodesize = 5, do.trace = 10, 
                            ntree=101, na.action=na.fail, importance=TRUE, proximity=FALSE)
      
   
    pointSub <- pointDat[pointDat$Zone == Z,] ###subset grid based on previously predicted zones
    pointSub$BGC<- predict(SZmodel, newdata = pointSub[,-c(1:4,22),]) ### predict subzones
    out <- pointSub[,c("ID1", "Latitude","Longitude","Zone", "BGC")]
    out
  }else{ ##if only one subzone, plot 
    pointSub <- pointDat[pointDat$Zone == Z,]
    #pointSub$Zone <- droplevels(pointSub$Zone)
    pointSub$BGC <- pointSub$Zone
    out <- pointSub[,c("ID1", "Latitude","Longitude","Zone", "BGC")]
    out
  }
  
}

grid.sbz <- dplyr::select(SZPred, ID1, BGC)#"Longitude","Latitude", "Elevation", "Zone")]
table(grid.sbz$BGC)
grid.sbz <- droplevels(grid.sbz)

###link to hex polygons
hexZone2 <- left_join(hexZone, grid.sbz, by = "ID1")
#temp <- select(hexZone, Zone, geometry)
temp <- hexZone2
temp$BGC <-  fct_explicit_na(temp$BGC , na_level = "(None)")
table(temp$BGC)

temp2 <- temp
st_precision(temp2) <- 0.5###union polygons within zsubone
t3 <- temp2 %>%
  group_by(BGC) %>%
  summarise(geometry = sf::st_union(geometry)) %>%
  ungroup()

mapview(t3)
t3 <- st_zm(t3, drop=T, what='ZM')
st_write(t3, dsn = "outputs", layer = "SubzoneRaw800m", driver = "ESRI Shapefile", update = TRUE)


```
#select a subset of training points by BGC from the final surface for full model build
library(clhs)
#BGC = "ICHxwz"
allUnits <- unique(pointLHC$Subzone)
pointLHC$ID1 <- row.names(pointLHC)
LHCtraining <- foreach(BGC = allUnits, .combine = rbind, .packages = c("clhs")) %dopar% {
temp <- pointLHC[(pointLHC$Subzone %in% BGC),]
Num <- countSubzone$sample[(countSubzone$Subzone %in% BGC)]
  samples <- clhs(temp[,-c(1:6)], 
                size = Num,           # Test a range of sample sizes
                iter = 100,        # Arbitrarily large number of iterations of optimization procedure. Default=10,000 but a larger number may be used
                progress = TRUE, 
                simple = FALSE)

cLHS <- samples$sampled_data
cLHS$ID1 <- row.names (cLHS)
cLHS_Points <- merge (pointLHC[,c(1:6)], cLHS, by = "ID1")
cLHS_Points
}

write.csv(LHCtraining, "USA_Training_LHC.csv")
X1 <- LHCtraining #feed these back into building a new rF model above

